Template for indicator documentation

### 2023-03-14 slitasje.Rmd
explain net steps and simplify workflow

10. Sample 1000 random points from an ecosystem delineation and extract
    indicator values and sd at those points.

11. Export indicator maps and regional extrapolated maps

@@

```{r, fig.cap="Slitasje indicator extrapolated over Trondheim"}
myCol <- "RdYlGn"
tmap_arrange(
tm_shape(wetlands_slitasje_extr_trd)+
  tm_polygons(col = "w_mean",
    title="Slitasje indicator",
    palette = myCol,
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="blue")+
  tm_layout(title = "Wetland"),

@@

Now, there are multiple way to proceed. We have

-   indicator maps (vector data, small polygons), and we have

-   total area for each HIA class and region combo, and we have the

-   extrapolated indicator maps (HIA maps with aggregated indicator
    values and errors, also vector data, wall-to-wall with some NA areas).

-   We also have, for wetlands, an ecosystem type (ET) map.

Option 1:

1.  crop (i.e. intersect) the extrapolated indicator maps to ecosystem
    delineation maps, and then;

2.  aggregate these for each region or accounting area.

Step 1 above requires an st_interactions on two very large vector data
sets with millions of polygons. This should be avoided.

Option 2:

1.  rasterize the extrapolated indicator map, using the ET map as a
    template, and

2.  mask it using the perfectly aligned ET map.

This might be faster.

Option 3:

@@

This is perhaps even faster again. I will try the last option, even
though it will not provide spatially explicit indicator values for the
entire ecosystem delineation, like option 2 does. Perhaps I will revisit
option 2 later.

@@

```{r, eval=F}
#takes about one minute without the proxy
mireET <- stars::read_stars("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Myrmodell/myrmodell90pros.tif",
                             proxy=T)
mireET
```

@@

So I will transform the extrapolated indicator map ti match the ET instead. Note that the ET map is just for southern Norway up to about the middle of the country.

```{r}
system.time(
  wetlands_slitasje_extr <- st_transform(wetlands_slitasje_extr, st_crs(mireET))
)
```

Then I import the real ET data and mask

```{r, eval=F}
mireET <- stars::read_stars("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Myrmodell/myrmodell90pros.tif",
                             proxy=F)
```

<!-- PLAN: rasterize wetland_slitasje_extr and c() with mireET, and mask -->
<!-- Or use  -->
<!-- wetlands_slitasje_extr[mireET==0] <- NA -->

```{r}
wetlands_slitasje_extr_stars <- stars::st_as_stars(wetlands_slitasje_extr)
temp <- c(wetlands_slitasje_extr_masked, mireET)
class(wetlands_slitasje_extr)
dim(temp)
```

@@

We need to aggreate the ET occurence and their indicator values, to a
regional delineation. This functionality will be developed in `eaTools`.
We need to produce distributions for the indicator valeus for each ET
occurrence (or groups of them) and resample the distributions wiht an
area weighting.

@@

But I will also export a map of the indicator where we extrapolate the
indicator values to the homogeneous areas. This map should be
interpreted with caution, since they seem to provide explicit indicator
values for areas where there is no data. This map is mainly for
aggregation purposes. I will mark this map with the prefix
*homogeneous-areas*. I can also export this same mappen after masking it
with the ET map, and this output must come with the same discalimers.

Finally I will export a map of the five accounting areas with
wall-to-wall indicator value and errors. This map can also contain data
covering other ETs. Otherwise the maps would not we visually
interpretative.

@@

Prepare export

>>>>

10. TO DO: Prepare ecosystem delineation maps and us ethese to mask the extrapolated indicator maps

11. Spatial aggregation of indicator values and uncertainties to accounting areas

12. Export indicator maps and regional extrapolated maps

@@

```{r extrapolated-slitasje, fig.cap="Slitasje indicator extrapolated over Trondheim"}
myCol <- "RdYlGn"
tmap_arrange(
tm_shape(wetlands_slitasje_extr_trd)+
  tm_polygons(col = "w_mean",
    title="Slitasje indicator",
    palette = myCol,
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="blue")+
  tm_layout(title = "Wetland"),

@@

The next step now is to

1. Prepare ecosystem delineation maps in EPSG25833 and perfectly aligned to a master grid

2. Rasterize the extrapolated indicator map, using the ET map as a template

3. and mask it using the perfectly aligned ET map.

A faster way could be to:

@@

... but his will not provide spatially explicit indicator values for the
entire ecosystem delineation.

@@

```{r, eval=T}
#takes about one minute without the proxy
(mireET <- stars::read_stars("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Myrmodell/myrmodell90pros.tif",
                             proxy=T))
```

@@

I will therefore ask GIS engineers to prepare ecosystem maps to the standard that we chose, so that we can simply combine the stars object with `stars::c()`, or perhaps side-step that operatio and just mask directly like this: `wetlands_slitasje_extr[mireET==0] <- NA`.

@@

We need to aggreate the ET occurence and their indicator values, to a
regional delineation. This functionality can be developed in `eaTools`,
but will in any case build around `exactextratr::exact_extract()`.

To carry the errors, we can for example
produce distributions for the indicator values for each ET
occurrence (or groups of them) and resample the distributions with an
area weighting.

@@

I can also export a map of the indicator where we extrapolate the
indicator values to the homogeneous areas (e.g. Fig. \@ref(fig:extrapolated-slitasje) . This map should be
interpreted with caution, since they appear as providing explicit indicator
values for areas where there is no data. This map is mainly for
aggregation purposes. I will mark this map with the prefix
*homogeneous-areas*. I will also export this same map after masking it
with the ET map, and this output must come with the same disclaimers.

Finally I will export a map of the five accounting areas with
wall-to-wall indicator value and errors (e.g. Fig \@ref(fig:wall-to-wall-summer-temp-indicator)). This map can also contain data
covering other ETs. Otherwise the maps would not we visually
interpretative.

@@

This is how I plan to export the final product.
